I first ran SWIG on the interface file.

```
swig -r beagle.i
```

I then compiled the wrapper library.

```
MAKEFLAGS="CFLAGS=-I/usr/local/include/libhmsbeagle-1" R CMD SHLIB beagle_wrap.c -L/usr/local/lib -lhmsbeagle
```

I renamed ```beagle.R``` to ```beagle_orig.R```, and search and replaced ", PACKAGE = 'beagle'", as the wrapper code isn't in a package at the moment, generating a new ```beagle.R```. I also had problems as "_p_int" was not a defined class. I simply commented out ```ans <- new("_p_int", ref=ans) ;``` in ```new_intArray```. I'm having problems with some of the array stuff in SWIG.

I now load the wrapper library and the R code.

```{r}
dyn.load(paste("beagle_wrap", .Platform$dynlib.ext, sep=""))
source("beagle.R")
```

Define the data.

```{r}
mars <- "CCGAG-AGCAGCAATGGAT-GAGGCATGGCG"
saturn  <- "GCGCGCAGCTGCTGTAGATGGAGGCATGACG"
jupiter <- "GCGCGCAGCAGCTGTGGATGGAAGGATGACG"
```

Get the number of site patterns.  These could optionally be compressed.

```{r}
nPatterns <- nchar(mars)
```

Create an instance of the BEAGLE library

```{r}
returnInfo <- BeagleInstanceDetails()
instance <- beagleCreateInstance(
        3,  	# Number of tip data elements (input)
				2,	        # Number of partials buffers to create (input) -- internal node count 
				3,		# Number of compact state representation buffers to create -- for use with setTipStates (input)
				4,		# Number of states in the continuous-time Markov chain (input) -- DNA
				nPatterns,	# Number of site patterns to be handled by the instance (input) -- not compressed in this case
				1,		# Number of eigen-decomposition buffers to allocate (input)
				4,		# Number of transition matrix buffers (input) -- one per edge
				1,		#  Number of rate categories
				0,		# Number of scaling buffers -- can be zero if scaling is not needed
				NULL,	 # List of potential resource on which this instance is allowed (input, NULL implies no restriction
				0,		# Length of resourceList list (input) -- not needed to use the default hardware config
				0,		# Bit-flags indicating preferred implementation charactertistics, see BeagleFlags (input)
				0,		# Bit-flags indicating required implementation characteristics, see BeagleFlags (input)
				returnInfo
)
```

Check that instance was created.

```{r}
if(instance<0){
  stop("Failed to obtain BEAGLE instance")
}
```


I define some utility functions that map bases to integers and fill in some arrays.

```{r}
getTable <- function(){
  l <- list()
  l["A"] <- 0
  l["C"] <- 1
  l["T"] <- 2
  l["G"] <- 3
  l["a"] <- 0
  l["c"] <- 1
  l["t"] <- 2
  l["g"] <- 3
  l["-"] <- 4
  l
}
createStates <- function(st,tab){
  states <- new_intArray(nchar(st))
  for(i in 1:nchar(st)){
    # Note zero-based indexing
    intArray_setitem(states,i-1,tab[[substr(st,i,i)]])
  }
  return(states)
}
```

```{r,eval=FALSE}
tab <- getTable()
marsStates <- createStates(mars,tab)
```

